% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
%\usepackage[review]{acl}
\usepackage[]{acl}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsfonts}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}

% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{mathtools}

\newcommand{\AL}[1]{{\color{blue}{[Andrew: #1]}}}

% If the title and author information does not fit in the area allocated, uncomment the following
%
%\setlength\titlebox{<dim>}
%
% and set <dim> to something 5cm or larger.

\title{Unsupervised Lexical Semantic Change Detection via Influence Functions}

% Author information can be set in various styles:
% For several authors from the same institution:
% \author{Author 1 \and ... \and Author n \\
%         Address line \\ ... \\ Address line}
% if the names do not fit well on one line use
%         Author 1 \\ {\bf Author 2} \\ ... \\ {\bf Author n} \\
% For authors from different institutions:
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \And  ... \And
%         Author n \\ Address line \\ ... \\ Address line}
% To start a seperate ``row'' of authors use \AND, as in
% \author{Author 1 \\ Address line \\  ... \\ Address line
%         \AND
%         Author 2 \\ Address line \\ ... \\ Address line \And
%         Author 3 \\ Address line \\ ... \\ Address line}

\author{
  Yuan-Hong Liao \\
  University of Toronto \\
  \texttt{andrew@cs.toronto.edu}}

\begin{document}
\maketitle
\begin{abstract}
  Languages change through time, and detecting the lexical semantic change in languages provides better understanding across different cultures, regions, and periods.
 Determining the ground truth of language semantic change requires high linguistics expertise. 
 Therefore, this work aims to detect the lexical semantic change between two corpora in an unsupervised manner.
 This work leverages the fact that language models learn word usage in a given corpus. 
 Once a word undergoes any semantic change, the language models trained on the corpus change accordingly.
 I implement this motivation via influence functions to efficiently approximate the model trained on different corpus configurations.
 In the SemEval2020 benchmark, the proposed approach achieves 0.7 accuracies in the binary task, outperforming 20 from 22 teams in the SemEval2020 challenges.
 In the ranking task, it achieves spearman's correlation of 0.59, outperforming all 22 teams in the SemEval2020 challenges.
\end{abstract}


\input{sec/intro.tex}
\input{sec/rw.tex}
\input{sec/method.tex}
\input{sec/exp.tex}
\input{sec/discussion.tex}

% Entries for the entire Anthology, followed by custom entries
\clearpage
\bibliography{anthology,custom}

%\appendix

%\section{Example Appendix}
%\label{sec:appendix}

%This is an appendix.

\end{document}
